---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# biascorrection

## Introduction
Collection of functions to de-bias daily ensemble forecasts. This package includes a variety of functions with a focus on ease-of-use rather than performance when applied to large datasets. Accordingly, specific functions will have to be implemented separately to be applicable in an operational setting.

The bias correction options include:

* mean de-biasing (`unbias`)
* mean de-biasing with loess smoothing (`smooth`, `smoothobs`)
* mean de-biasing with variance scaling (`smooth_scale`, `smoothobs_scale`)
* monthly-mean de-biasing (`monthly`)
* climate conserving recalibration based (`ccr`) based on smoothed climatology (`smoothccr`) or monthly climatology (`ccr_monthly`)
* quantile mapping (`qqmap`)
* mean de-biasing with linear time trend (`trend`)
* mean de-biasing with bias conditional on signal (`conditional`)
* ...

## Getting started

First, install the package and vignettes.

```{r, echo=TRUE, eval=FALSE}
devtools::install_github("jonasbhend/biascorrection", build_vignettes=TRUE)
```

Next, check out the examples provided in the vignette or the `help` and `examples` of individual functions in `biascorrection`.

```{r, echo=TRUE, eval=FALSE}
vignette('biascorrection')
```